A generalization of Plackett-Luce likelihoods to incorporate non-independence of competitors

### 2023-01-06 inst/plackett_luce_monster.Rmd
better (start simple!) examples

@@

Might be better to consider ${5\choose 2\,3}$ arrangements of $aaabb\ldots bbaaa$.

>>>>

# Start simple

The simplest nontrivial case is three runners, $\left\lbrace
1,2,3\right\rbrace$ where runner 1 helps runner 2 [and that is the
only non PL interaction].  Note that this is _not_ a symmetric
interaction: 1 helps 2, but 2 does not help 1.  The model is that 1
helps 2 if and only if 1 has finished and 2 has not.

123: $\frac{p_1}{p_1+p_2+p_3}\cdot\frac{\lambda p_2}{\lambda p_2+p_3}$

132: $\frac{p_1}{p_1+p_2+p_3}\cdot\frac{p_3}{\lambda p_2+p_3}$

213: $\frac{p_2}{p_1+p_2+p_3}\cdot\frac{p_1}{p_1+p_3}$

231: $\frac{p_2}{p_1+p_2+p_3}\cdot\frac{p_1}{p_1+p_3}$

312: $\frac{p_3}{p_1+p_2+p_3}\cdot\frac{p_1}{p_1+p_2}$

321: $\frac{p_3}{p_1+p_2+p_3}\cdot\frac{p_2}{p_1+p_2}$

(Above, the probabilities clearly sum to 1).  Now the same but with
four runners $\left\lbrace 1,2,3\right\rbrace$, again with runner 1
helping runner 2.  Thus, in a situation where $p_1$ has finished and
$p_2$ is still running, we have $p_2\longrightarrow\lambda p_2$
[although the $\lambda$ cancels if 2 is the only competitor still
running, that is, if *all* the remaining competitors are helped]

1234: $\frac{p_1}{p_1+p_2+p_3+p_4}\cdot\frac{\lambda p_2}{\lambda p_2+p_3+p_4}\cdot\frac{p_3}{p_3+p_4}$

1243: $\frac{p_1}{p_1+p_2+p_3+p_4}\cdot\frac{\lambda p_2}{\lambda p_2+p_3+p_4}\cdot\frac{p_4}{p_3+p_4}$

1324: $\frac{p_1}{p_1+p_2+p_3+p_4}\cdot\frac{p_3}{\lambda p_2+p_3+p_4}\cdot\frac{\lambda p_2}{\lambda p_2+p_4}$

1342: $\frac{p_1}{p_1+p_2+p_3+p_4}\cdot\frac{p_3}{\lambda p_2+p_3+p_4}\cdot\frac{p_4}{\lambda p_2+p_4}$

1324: $\frac{p_1}{p_1+p_2+p_3+p_4}\cdot\frac{p_3}{\lambda p_2+p_3+p_4}\cdot\frac{\lambda p_2}{\lambda p_2+p_4}$

1423: $\frac{p_1}{p_1+p_2+p_3+p_4}\cdot\frac{p_4}{p_1+\lambda p_2+p_3+p_4}\cdot\frac{\lambda p_2}{\lambda p_2+p_3}$

1432: $\frac{p_1}{p_1+p_2+p_3+p_4}\cdot\frac{p_4}{p_1+\lambda p_2+p_3+p_4}\cdot\frac{p_3}{\lambda p_2+p_3}$

2xxx: P-L [no help given by 2 to 1]

3124: $\frac{p_3}{p_1+p_2+p_3+p_4}\cdot\frac{p_1}{p_1+p_2+p_4}\cdot\frac{\lambda p_2}{\lambda p_2+p_4}$

3142: $\frac{p_3}{p_1+p_2+p_3+p_4}\cdot\frac{p_1}{p_1+p_2+p_4}\cdot\frac{p_4}{\lambda p_2+p_4}$

32xx: P-L

3412: P-L [1 helps 2 but when 1 crosses the finishing line, only 2 is
still running; there are no non-helped competitors]

4123: $\frac{p_4}{p_1+p_2+p_3+p_4}\cdot\frac{p_1}{p_1+p_2+p_3}\cdot\frac{\lambda p_2}{\lambda p_2+p_3}$

4132: $\frac{p_4}{p_1+p_2+p_3+p_4}\cdot\frac{p_1}{p_1+p_2+p_3}\cdot\frac{p_3}{\lambda p_2+p_3}$

42xx: P-L

43xx: P-L

# Next simplest: 1 and 2 mutually support each other:

123: $\frac{p_1}{p_1+p_2+p_3}\cdot\frac{\lambda 2}{\lambda p_2+p_3}$

132: $\frac{p_1}{p_1+p_2+p_3}\cdot\frac{p_3}{\lambda p_2+p_3}$

213: $\frac{p_2}{p_1+p_2+p_3}\cdot\frac{\lambda p_1}{p_1+p_3}$

231: $\frac{p_2}{p_1+p_2+p_3}\cdot\frac{p_3}{\lambda p_1+p_3}$

312: $\frac{p_3}{p_1+p_2+p_3}\cdot\frac{p_1}{p_1+p_2}$

321: $\frac{p_3}{p_1+p_2+p_3}\cdot\frac{p_2}{p_1+p_2}$

# Four runners

@@

Might be better to consider ${5\choose 2\,3}$ arrangements of
$aaabb\ldots bbaaa$.  FOLLOWING IS NOT RIGHT: THE PROBABILITY MODEL IS
NOT PRECISELY DEFINED.
